---
title: "Factors affecting Sleep"
author: "540214264, 540754276, 540062713, 540286434, 540301162, 540496066"
subtitle: "Project 2"
date: "University of Sydney | DATA1001 | APRIL 2024"

output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    css: bootstrap.min.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    

---

<br>

# Executive Summary

-   The aim of this report is to ascertain the main factor causing sleep deprivation among young adults today
-   Data collected shows a positive correlation between screen time and quality of sleep 
-   Those who use their phones before bedtime tend to suffer from worse sleeping quality compared to those who don't 

<br>



# Full Report

<br>

## Initial Data Analysis (IDA)



### Source

The data used in this report was collected from university students from 3/28/24 to 4/14/24 through the usage of Google Forms:
(https://forms.gle/xxXj6JNymotvkEot5)



### Structure

We received 30 responses for our survey. Each row represents a survey response while each of the columns represent our variables. This survey contained both quantitative (Hours of sleep, Screen time, Phone before bed) and qualitative (Exercise, Time of medication) variables. 

```{r, warning=FALSE, message=FALSE}
# Shows barplot of ages of survey participants
library(tidyverse)
library(readr)
Responses = read_csv("./Project 2 responses - Sheet1.csv", show_col_types = FALSE)
Responses = filter(Responses, Age < 30)
ggplot(Responses) + 
  geom_histogram(aes(x = Age), bins=30, fill = "steelblue") +
  labs(y = "Count") +
  scale_x_continuous(breaks = 0:23) 

```

```{r, warning=FALSE, message=FALSE}
#shows responses and count
age_labels <- c("17", "18", "19", "20", "23")
response_counts <- c(2, 13, 9, 4, 1)
cat("Age: ", paste(age_labels, collapse = " "), "\n")
cat("Responses: ", paste(response_counts, collapse = " "), "\n")
```

### Limitations

Not everyone has a fixed sleeping schedule. It is quite likely that our survey participants had fluctuating hours of sleep throughout their weeks. This could be due to their differing school timetables, causing them to sleep and wake up at varying times.Also, the sleep quality is biased on how people judge their own sleep quality, because we are aware that everyone has their own perspective about what a good sleep quality is. Furthermore, we can see from the graph that most of our responses are from those who are 18 or 19. There is also a very small range of ages from 17 - 23 years with no responses from ages 21 and 22.

### Assumptions
We assume that all survey participants will tell the truth and will take the survey seriously. Any illogical responses (i.e. age of participant being 999 years old) will not be included in our results and will be assumed to be misinformation, despite lack of evidence.

### Data Clearing
For data clearing, we will use a function called `filter()` to remove any illogical results that may damage the reliability of our results. Responses with age being older than 30 were excluded, as we are targeting young adults 30 and below. Values that are not numerical under 'caffeine per day' has been set as non applicable data.


<br>
<br>
<h2> Research theme </h2>
We will investigate the relationship between the sleeping behavior of University students and the activities they partake in during their normal day.


## Research Question 1

How does phone usage before bed affects one's sleep quality and hours?

```{r, warning=FALSE, message=FALSE}
age_labels <- c("17", "18", "19", "20", "23")
response_counts <- c(2, 13, 9, 4, 1)
cat("Age: ", paste(age_labels, collapse = " "), "\n")
cat("Responses: ", paste(response_counts, collapse = " "), "\n")
```



```{r message = FALSE, warning = F}
library(tidyverse)
ggplot(Responses, aes(x = `Quality of sleep`)) + 
  geom_bar(aes(fill= `Phone before bed`)) +
  ggtitle("Quality of Sleep ") +
  labs(x = "Quality of Sleep", y = "Frequency", fill = "Phone usage before bed") +
  scale_y_continuous(breaks = 0:15) +
  guides(fill = guide_legend(reverse = T))
```


Those who use their phones before bed tend to have lower sleep quality. We assume those who still have good sleep quality despite using phone before bed are special case ( they may exercise regularly, thus making them having better sleep quality ). The average of those who don't use a phone before bed have better quality of sleep compared to those who do. 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
sleep_time_levels <- c("20:00:00", "20:30:00", "21:00:00", "21:30:00", 
                       "22:00:00","22:30:00","23:00:00","23:30:00",
                       "00:00:00","00:30:00","01:00:00", "01:30:00",
                       "02:00:00", "10:00:00", "11:00:00")


ggplot(Responses) +
  geom_bar(aes(factor(`Sleep time`, sleep_time_levels), fill = `Phone before bed`)) +
  scale_y_continuous(breaks = 1:10) +
  ggtitle("Usage of phone VS Sleeping times") +
  ylab("Frequency") +
  xlab("Sleeping Times") +
  coord_flip()

```


Looking at the graph, Those who don't use their phones before bedtime go to bed at earlier times. Most people who use their phones seem to go to bed at 1AM . On the other hand, those who don't use their phones go to bed earlier with only 1 person sleeping at 1AM, the rest sleeping earlier. However, we also have to take into account the lack of participants who don't use their phone before bed.
### Conclusion:
As evident through the results, young adults who use their phones before bedtime are more likely to go to bed at later times as well as experience poorer quality of sleep. Those who don't, are more likely to sleep at earlier times and experience better sleep quality.


## Research Question 2

How does screen time influence duration of sleep?

```{r, message = F, warning = F}
Responses$`Hours of sleep` <- as.numeric(Responses$`Hours of sleep`)
Responses$`Screen time` <- as.numeric(Responses$`Screen time`)
ggplot(Responses, aes(x=`Screen time`, y=`Hours of sleep`)) +
  geom_point() + 
  geom_smooth(method = 'lm', se = T) +
  scale_x_continuous(breaks = 0:16) + 
  labs(x = "Screen Time (Hours)", y = "Sleeping Duration (Hours)", title = "Scatter plot of Sleeping Duration against Screen Time")
```
```{r, message = F, warning = F}
correlation = cor(Responses$`Screen time`, Responses$`Hours of sleep`)
print(paste0("Correlation Coefficient = ", correlation))
```

<br>
From this linear regression graph, we can conclude that sleeping duration is not influenced by screen time since the correlation coefficient shown above is near 0 and positive. However, the area of the standard error above is incredibly significant. Hence, it could also be said that the projected values shown above are far from accurate as well.

```{r}
l_m = lm(`Hours of sleep` ~ `Screen time`, Responses)
ggplot(l_m, aes(x=.fitted, y=.resid)) +
  geom_point() +
  ylab("Residual") +
  xlab("Fitted") +
  ggtitle("Residual vs Fitted") +
  geom_hline(yintercept = 0, color="blue")
```


From the residual plot above, we can see that the residual plot appears to be random and supports the idea that the linear fit is appropriate for the data. Showing a lack of correlation between sleeping duration and screen time, suggesting that sleep duration is not affected by screen during the day.


Summary:<br>
The data we have received is insufficient to make a hypothesis as to whether screen time has an affect on duration of sleep.

## Article
Studies exploring the relationship between screen time and quality corresponded with data collected to an extent; with Asamaa's study (2023) observing that people who used smartphones before bedtime experienced significantly worse sleep quality with 64.2% reporting issues, compared to only 31.7% of non-users. Furthermore, smartphone users at bedtime also reported to have shorter durations of sleep.


# References

Style: APA
<br>
Asmaa Abdelghany Elsheikh, Safaa Abdelfattah Elsharkawy, & Doaa Sadek Ahmed. (2023). Impact of smartphone use at bedtime on sleep quality and academic activities among medical students at Al -Azhar University at Cairo. Journal of Public Health. https://doi.org/10.1007/s10389-023-01964-8

‌
<br>

# Acknowledgment
24/2/2024 Online |Braxton, William, Vincent, Alex, Mingye|- Figuring out what we want to explore and formulating research question
<br>
2/3/2024 Online |Braxton, William, Vincent, Mingye|- Started to list out variables and formulating questions
<br>
3/3/2024 Online |Braxton, William, Vincent, Mingye| - Finished off Project 1 and created a survey in Google forms
<br>
10/03/2024 Online on discord |Braxton, William, Vincent, Damara| - Designing the plan and research topic for Project plan
<br>
14/03/2024 After Lab |Braxton, William, Vincent, Alex, Damara, Mingye| - Creating the survey questions and the google form
<br>
28/03/2024 After Lab |Braxton, William, Mingye, Damara| - Posting the survey to obtain results
<br>
15/04/2024 Online on Github |Braxton| - Creating github repo and initial scaffolding for report
<br>
17/04/2024 Online on Github |Damara| - Added research questions
<br>
18/04/2024 After Lab Online |Braxton, William, Vincent, Alex, Mingye, Damara| - Communicating report data 
<br>
18/04/2024 Online on Github and Discord | Braxton, William, Vincent, Damara| - Writing out report 
